---
title: "Homework 7: Data Transformations"
author: "Key"
format: html
---

This week's assignment uses data from Tidy Tuesday ([link](https://github.com/rfordatascience/tidytuesday/tree/2e9bd5a67e09b14d01f616b00f7f7e0931515d24/data/2020/2020-02-18)) and relates to food consumption and CO2 emissions. 

```{r data-code, include = F}
# Credit to Kasia and minorly edited to create output file and test plot
# Blog post at https://r-tastic.co.uk/post/from-messy-to-tidy/
library(rvest)
library(dplyr)

url <- "https://www.nu3.de/blogs/nutrition/food-carbon-footprint-index-2018"

# scrape the website
url_html <- read_html(url)

# extract the HTML table
whole_table <- url_html %>% 
  html_nodes('table') %>%
  html_table(fill = TRUE) %>%
  .[[1]]

table_content <- whole_table %>%
  select(-X1) %>% # remove redundant column
  filter(!dplyr::row_number() %in% 1:3) # remove redundant rows

raw_headers <- url_html %>%
  html_nodes(".thead-icon") %>%
  html_attr('title')

tidy_bottom_header <- raw_headers[28:length(raw_headers)]
# tidy_bottom_header[1:10]

raw_middle_header <- raw_headers[17:27]
# raw_middle_header

tidy_headers <- c(
  rep(raw_middle_header[1:7], each = 2),
  "animal_total",
  rep(raw_middle_header[8:length(raw_middle_header)], each = 2),
  "non_animal_total",
  "country_total")

# tidy_headers

combined_colnames <- paste(tidy_headers, tidy_bottom_header, sep = ';')
colnames(table_content) <- c("Country", combined_colnames)

table_content <- table_content %>%
  mutate_at(vars(2:26), as.numeric)
```

```{python, include = F}
table_content = r.table_content # Grab data from R for use in python
```

The code above reads the data in from the [original webpage](https://www.nu3.de/blogs/nutrition/food-carbon-footprint-index-2018) and gets it into tabular form. 

Your job is to complete the following tasks:

1. Describe the state of the data set, `table_content`. 
    - What are the variables in the data set?
        - Country
        - Food
        - Consumption
        - Carbon Output
    - Is it in tidy form? What principles of tidy data does this violate?    
    No. Each variable does not have its own column (food is spread out over multiple columns). Each observation does not have its own row (multiple observations are in a single row).
    - What steps do you need to take to get it into tidy form?
        1. Convert to ultra-long format so that each food type has its own row (consumption and carbon output will be together in a single column)
        2. Split food type from unit so that they are in separate columns.
        3. Convert to wider format, separating consumption from carbon output so that each is in its own column.

2. Sketch out what the final (tidy) data set will look like. You can use markdown table syntax or a picture here, but if you use a picture, upload it to imgur and include the image link in this document USING PROPER MARKDOWN SYNTAX. 

Country | Food | Consumption | Carbon
--- | --- | --- | ---
Argentina | Pork | 10.51 | 37.20
Australia | Pork | 24.14 | 85.44
... | ... | ... | ...
Argentina | Chicken | 38.66 | 41.53
Australia | Chicken | 46.12 | 49.54
... | ... | ... | ...


3. Write R or python code for each step in the process you identified in #1. Show what the data looks like at each step using `head()`. Each step should be in a different code chunk. 

Convert to ultra-long format so that each food type has its own row

```{r}
library(tidyr)
step1 <- table_content %>%
  pivot_longer(-Country, names_to = "variable", values_to = "value")
head(step1)
```

Split food type from unit so that they are in separate columns.

```{r}
step2 <- step1 %>%
  separate(variable, into = c("Food", "Unit"), sep = ";")
head(step2)
```

Convert to wider format, separating consumption from carbon output so that each is in its own column.

```{r}
step3 <- step2 %>%
  pivot_wider(names_from = Unit, values_from = value)
head(step3) 
```

Clean up names

```{r}
step4 <- step3 %>%
  rename(Consumption = `Supplied for Consumption (kg/person/year)`, Carbon = `Kg CO2/person/year`)
head(step4)
```


4. For each food type (you may have to remove total values), plot the relationship between Carbon output and Consumption (use facets to get separate plots for each type of food). What do you notice for each plot? If you want to reduce carbon emissions, what foods should you eat less of? 

```{r}
library(ggplot2)
step4 %>%
  filter(!grepl("total", Food)) %>%
ggplot(aes(x = Consumption, y = Carbon, color = Food)) + geom_point() + facet_wrap(~Food)

```

The data shows that the consumption of beef/lamb/goats is associated with significantly higher carbon output per capita, so eating less of those products would produce lower overall carbon emissions.

5. Look at the plot above again. Do you have any concerns about the data? The data source? 

The points are all in a straight line. This suggests that the data we have been working with are not collected observations (there would be random noise) but rather calculated based on a constant rate of emissions per unit of consumption. Reading the information at the original website (https://www.nu3.de/blogs/nutrition/food-carbon-footprint-index-2018), we can see that the data source 
is promoting veganism and is calculating carbon footprint per capita for each food type based on the amount consumed in the country and the country's total CO2 produced. This is ridiculous for a number of reasons - global trade among them - but it also neglects different agricultural practices that may exist in raising e.g. corn-fed beef vs. grass-fed beef, manufacturing that is a driver of CO2 emissions unrelated to agriculture, and many other factors. 
